Title

### 2024-05-13 paper/main.tex
Updates

\begin{abstract}
	\emph{Causal Discovery} involves learning causal structures from data,
	and has been studied both in Directed Acyclic Graphs (DAGs) and
	Structural Equation Models (SEMs). This problem has been extensively
	studied for both DAGs and SEMs. In the field of DAGs, many automated
	algorithms have been developed for constructing a DAG from a given
	dataset. However, the adaption of these algorithms in applied fields
	have been limited, possibly due to a lack of confidence in these
	algorithms stemming from them making obvious mistakes, and the
	difficulties in choosing the right algorithm for a given dataset.
	Consequently, in applied disciplines, these DAGs are predominantly
	constructed manually based on domain expertise. Due to being
	constructed by hand, it is vital that the fit of these models are
	tested against data and the models are potentially modified to acheive
	a better fit. On the other hand, in the field of SEMs models are
	typically constructed manually through an iterative process of
	assessing fit and modifying the model according to that. This process
	is commonly known as \emph{Specification Search} where methods like
	\emph{modification indices} can be used to guide this modification
	process. In the case of DAGs, although there are tests such as Shipley
	2000 that can assess the global fit of model by combining the tests
	using implied conditional independencies of the model, but there are no
	methods that can guide our modification process. In this paper, we
	present a simple modification process aimed at helping researchers in
	manually constructing or modifying the DAGs outputted by an algorithm.
	This modification process is based on using a measure of association
	for these individual local CI tests to identify the worst offending CI
	test and combining it with expert knowledge. We also present a measure
	of association based on canonical correlations for mixed data. We also
	present a graphical web-tool that can assist researchers in iteratively
	modifying and constructing their DAGs. We theoretically show that in
	the presence of an oracle, this iterative modification process is
	consistent in recovering the true DAG. Emprically, we show that this
	process performs comparably to PC and Hill-Climb Search algorithms if
	the user is able to correctly identify the direction of the edge in one
	out of three cases.
\end{abstract}

@@

To perform any causal effect identification or estimation in Pearl's Directed
Acyclic Graphs (DAGs) based causal framework usually requires to come up with
the DAG structure. Learning these DAG structures from data is known as
\emph{causal discovery}. Plenty of algorithms have been developed for these in
the literature such as constraint-based methods such as PC, FCI, score based
methods such as Hill-Climb Search, GES, and constrained continuous optimization
based such as No TEARS. All these algorithms take different approaches to
causal discovery, constraint-based algorithms utilize statistical conditional
independence (CI) tests to come up with a model structure, the score-based
methods iteratively perform local modifications to the model such that some
score function is maximized, the continuous optimization methods reformulate
the problem statement as a constrained optimization problem and use
optimization algorithms to come up with a graph.

@@

Both in the cases when the DAG is an output of an algorithm or when constructed
by hand, it is vital to test these models against data before using them for
making any inference.

In the field of Structural Equation Models (SEMs), models are typically
constructed by hand. Researchers typically come up with an initial model, which
then they use the data to make modifications to. This process is known as
\emph{Specification Search}. One of the common ways to do this is through
\emph{modification indices} where a global fit of the model is measured with
different modification to the model. Utilizing expert knowledge to select the
appropriate modification that gives the maximum improvement to the global fit.

In the case of DAGs however, we do not have any way to As these CI based
testing uses p-values and effect size measures from CI tests for testing. And
as mixed data is common, we need both a test and effect size measure for mixed
data. Recently, a few CI tests for mixed data have been proposed but an effect
size is still missing.

\begin{enumerate}
	\item We present a procedure to construct and modify DAGs using domain
		knowledge.
	\item The presented procedure uses a (conditional) measure of
		association between variables. In the simple case of continuous
		or discrete variables, effect size measures can be used,
		however as no such measure exists for mixed data, we present
		one.
	\item Lastly, to assist researchers in using the DAG construction
		approach, we developed a web-based tool to help researches in
		constructing and modifying their DAGs.
\end{enumerate}

>>>>

\begin{abstract}
	\emph{Causal Discovery} involves learning causal structures from data,
	and has been studied both in Directed Acyclic Graphs (DAGs) and
	Structural Equation Models (SEMs). This problem has been extensively
	studied for both DAGs and SEMs. In the field of DAGs, many automated
	algorithms have been developed for constructing a DAG from a given
	dataset. However, the adaption of these algorithms in applied fields
	have been limited, possibly due to a lack of confidence in these
	algorithms stemming from them making obvious mistakes, and the
	difficulties in choosing the right algorithm for a given dataset.
	Consequently, in applied disciplines, these DAGs are predominantly
	constructed manually based on domain expertise. Due to being
	constructed by hand, it is vital that the fit of these models are
	tested against data and the models are potentially modified to acheive
	a better fit. On the other hand, in the field of SEMs models are
	typically constructed manually through an iterative process of
	assessing fit and modifying the model according to that. This process
	is commonly known as \emph{Specification Search} where methods like
	\emph{modification indices} combined with expert knowledge can be used
	to guide this modification process. In the case of DAGs, although there
	are tests to assess the global fit of a model by combining the tests
	using implied conditional independencies of the model, but there are no
	methods that can guide our modification process. In this paper, we
	present a simple modification process aimed at helping researchers in
	manually constructing or modifying DAGs outputted by an algorithm. This
	modification process is based on using a (conditional) measure of
	association between variables in the model to assess potential edges
	that best explains unexplained correlation between variables which the
	researcher specifies the direction of this edge. In the case of
	continuous or discrete variables, the effect size measures of local
	tests can be used as this measure of association, however for mixed
	data there is no such measure. We present a measure of association
	based on canonical correlations for mixed data. We also present a
	graphical web-tool that can assist researchers in iteratively modifying
	and constructing their DAGs. We theoretically show that in the presence
	of an oracle, this iterative modification process is consistent in
	recovering the true DAG. Emprically, we show that using the mixed data
	measure of association, the modification process performs comparably to
	PC and Hill-Climb Search algorithms if the user is able to correctly
	identify the direction of the edge in one out of three cases.
\end{abstract}

@@

To perform any causal effect identification or estimation using Directed
Acyclic Graphs (DAGs) or Structural Equation Models (SEMs), researchers need to
first construct a causal structure. This process of constructing the casual
structure from data is known as \emph{Causal Discovery}. In the field of DAGs,
plenty of algorithms have been developed to automatically construct DAGs from
data. These algorithm takes different approaches to causal discovery such as
Constraint-based algorithms such as PC and Fast Causal Infenrece (FCI) where
the algorithm tries to match the conditional independences in the data to the
ones impled by the model, score-based methods such as Hill-Climb Search, Greedy
Equivalence Search (GES)  where the algorithm tries to find a model that has
the best score given some scoring metric, and continuous optimization where the
problem is formulated as a constrained optimization problem such as No Tears.

@@

Both in the cases when the DAG is an output of an algorithm or when constructed
by hand, it is vital to test these models against data and potentially modifying 
them for a better fit before using them for making any inference. Both global
and local tests for testing the fit of these models are available. However, when
the fit fails, we don't have a procedure to modify these models.

In the field of Structural Equation Models (SEMs), models are also typically
constructed by hand. Researchers typically come up with an initial model, which
then they use the data and expert knowledge to make modifications to. This
process is known as \emph{Specification Search}. One of the common ways to do
this is through \emph{modification indices} where a global fit of the model is
measured with different modification to the model. Utilizing expert knowledge
to select the appropriate modification that gives the maximum improvement to
the global fit.

We present a similar procedure for modifying DAGs based on expert knowledge and
measure of association. We first use the measure of association to test whether
the association observed in the data is explained by the model structure. Based 
on this testing, we suggest adding edges by ranking them by the most unexplained
association between variables. The researcher can then use expert knowledge to 
decide the direction of the edge between the variables.

Our main contributions in this papers are as follows:
\begin{enumerate}
	\item We present a measure of (conditional) assocation between mixed variables using
		canonical correlations. This mixed variable association is a generalization 
		of partial correlation to mixed variables.
	\item We present a procedure to iteratively construct and modify DAGs using expert 
		knowledge such that it improves the explained association between variables.
	\item Lastly, to assist researchers in manully constructing and modifying DAGs,
		we developed a web-based tool to iteratively modify their DAGs.
\end{enumerate}
